---
title: "Hierarchy Experiment: Statistical model"
author: "Bodo Winter (statistical analysis)"
date: "7/17/2019"
output: html_document
---

## Introduction

This is the statistical analysis of our "hierarchy experiment", a replication-extension of Schubert's (2005) study which associates vertical positions with power. The present study includes the added dimension of gender.

This script takes the fitted Bayesian models and interprets them:

## Preprocessing

Load stuff:

```{r, message = FALSE}
# Packages:

library(tidyverse)
library(brms)
library(ggrepel)

# ggplot2 theme:

source('timo_theme.R')
```

## Load the model:

Load the main model that is fitted by the file "bayes_model_estimation.R":

```{r}
load('main_mdl.RData')
```

## Summarize the model:

```{r}
summary(main_mdl)
```

Reminder of deviation coding scheme:

* subject gender: female = -0.5, male = +0.5
* task: powerless = -0.5, powerful = +0.5
* hierarchy: powerful + down = -0.5, powerful + up = +0.5
* gender: male + down = -0.5, male + up = +0.5

## Posterior probabilities:

Calculate the posterior probabilities of the respective effects. First, extract posteriors:

```{r}
p <- posterior_samples(main_mdl)

subjGender <- p$'b_SubjGender1'
task <- p$'b_Task1'
hierarchy <- p$'b_Hierarchy1'
gender <- p$'b_Gender1'
hierarchy_gender <- p$'b_Hierarchy1:Gender1'
subjGender_hierarchy <- p$'b_SubjGender1:Hierarchy1'
subjGender_Gender <- p$'b_SubjGender1:Gender1'
task_gender <- p$'b_Task1:Gender1'
task_hierarchy <- p$'b_Task1:Hierarchy1'
threeway <- p$'b_SubjGender1:Hierarchy1:Gender1'
```

Calculate posterior probabilities:

```{r}
# Subject effect:

sum(subjGender < 0) / nrow(p)

# Task effect:

sum(task < 0) / nrow(p)

# Gender effect:

sum(gender < 0) / nrow(p)

# Hierarchy effect:

sum(hierarchy < 0) / nrow(p)

# Hierarchy * task interaction:

sum(task_hierarchy < 0) / nrow(p)

# Task * gender interaction:

sum(task_gender < 0) / nrow(p)

# Hierarchy * gender interaction:

sum(hierarchy_gender < 0) / nrow(p)

# Subject gender * hierarchy interaction:

sum(subjGender_hierarchy< 0) / nrow(p)

# Subject gender * gender interaction:

sum(subjGender_Gender < 0) / nrow(p)

# Threeway interaction:

sum(threeway < 0) / nrow(p)
```

## Plot the marginal effects:

Plot the marginal effects, task:

```{r, fig.width = 8, fig.height = 6}

marginal_effects(main_mdl, 'Task')
```

Plot the marginal effects, gender:

```{r, fig.width = 8, fig.height = 6}

marginal_effects(main_mdl, 'Gender')
```

Plot the marginal effects, hierarchy:

```{r, fig.width = 8, fig.height = 6}

marginal_effects(main_mdl, 'Hierarchy')
```

Hierarchy * gender interaction:

```{r, fig.width = 8, fig.height = 6}

marginal_effects(main_mdl, 'Gender:Hierarchy')
```

Task * hierearchy interaction:

```{r, fig.width = 8, fig.height = 6}
task_hierarchy <- marginal_effects(main_mdl, 'Task:Hierarchy')
plot(task_hierarchy)
```

Three-way interaction:

```{r, fig.width = 8, fig.height = 6}

conditions <- tibble(SubjGender = factor(c('female', 'male')))

interaction_preds <- marginal_effects(main_mdl, 'Gender:Hierarchy',
                 conditions = conditions)

plot(interaction_preds)
```

## Publication-ready plot of the marginal effects:

Extract the relevant data frames:

```{r}
task_hierarchy <- task_hierarchy$`Task:Hierarchy`
interaction_preds <- interaction_preds$`Gender:Hierarchy`
```

Plot the task * hierarchy results:

```{r}
p1 <- task_hierarchy %>%
  mutate(Hierarchy = ifelse(Hierarchy == 'BottomPower', 'power = down', 'power = up'),
         Task = ifelse(Task == 'powerful', 'more powerful?', 'less powerful?')) %>% 
  ggplot(aes(x = Hierarchy, col = Hierarchy, y = estimate__))
p1 + geom_errorbar(aes(ymin = lower__, ymax = upper__), size = 1.25, width = 0.1) + 
  geom_point(size = 6, pch = 15) +
  facet_wrap(~Task) + 
  theme_bham_stats +
  labs(title = 'Task * Hierarchy Condition interaction\n') +
  ylab('Predicted log response time\n') +
  xlab('\nHierarchy Condition') +
  theme(plot.title = element_text(face = 'bold', size = 20, hjust = 0.5),
        axis.title = element_text(face = 'bold', size = 20),
        axis.text.x = element_text(face = 'bold', size = 16),
        axis.text.y = element_text(face = 'bold', size = 12),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        legend.position = 'top') +
  scale_color_viridis_d(option = 'D', direction = -1, begin = 0.1, end = 0.7)
ggsave('../figures/figure2.png', width = 10, height = 7)
```

Plot the subject / hierarchy / gender interaction:

```{r}
p2 <- interaction_preds %>%
  mutate(Hierarchy = ifelse(Hierarchy == 'BottomPower', 'power = down', 'power = up'),
         Task = ifelse(Task == 'powerful', 'more powerful?', 'less powerful?'),
         Gender = ifelse(Gender == 'FemaleTop', 'female = up', 'male = up')) %>% 
  ggplot(aes(x = Gender, col = Hierarchy, y = estimate__))
p2 + geom_errorbar(aes(ymin = lower__, ymax = upper__), size = 1.25, width = 0.1,
                   position = position_dodge(width = 0.3)) + 
  geom_point(size = 6, pch = 15, position = position_dodge(width = 0.3)) +
  facet_wrap(~SubjGender) + 
  theme_bham_stats +
  labs(title = 'Gender Condition * Hierarchy Coondition * Participant Gender interaction\n') +
  ylab('Predicted log response time\n') +
  xlab('\nGender Condition') +
  theme(plot.title = element_text(face = 'bold', size = 20, hjust = 0.5),
        axis.title = element_text(face = 'bold', size = 20),
        axis.text.x = element_text(face = 'bold', size = 16),
        axis.text.y = element_text(face = 'bold', size = 12),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        legend.position = 'top') +
  scale_color_viridis_d(option = 'D', direction = -1, begin = 0.1, end = 0.7)
ggsave('../figures/figure3.png', width = 10, height = 7)
```



## Extract the random effects coefficients for items:

Get the PairName df out this object with all the estimates:

```{r}
pairs <- coef(main_mdl)$PairName
pairs <- pairs[, 1, 1:11]
head(pairs)
```

Extract the relevant random effects estimates:

```{r}
gender_coefs <- pairs[, 'Gender1']
hierarchy_coefs <- pairs[, 'Hierarchy1']
gender_interact_coefs <- pairs[, 'Hierarchy1:Gender1']
gender_interact_coefs <- pairs[, 'Hierarchy1:Gender1']
```

Make them into tidy data frames:

```{r}
gender_coefs <- tibble(GenderCoef = gender_coefs,
                       Pair = names(gender_coefs))
hierarchy_coefs <- tibble(HierarchyCoef = hierarchy_coefs,
                          Pair = names(hierarchy_coefs))
interact_coefs <- tibble(InteractionCoef = gender_interact_coefs,
                         Pair = names(gender_interact_coefs))
```

Plot them. First the gender coefficients:

```{r, fig.width = 12, fig.height = 6}
gender_coefs %>% ggplot(aes(x = reorder(Pair, GenderCoef),
                            y = GenderCoef)) +
  geom_col(fill = 'peachpuff3') + theme_bham_stats +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = 'By-item gender coefficients') +
  xlab('')
```

Plot them. Next the hierarchy coefficients:

```{r, fig.width = 12, fig.height = 6}
hierarchy_coefs %>% ggplot(aes(x = reorder(Pair, HierarchyCoef),
                            y = HierarchyCoef)) +
  geom_col(fill = 'peachpuff3') + theme_bham_stats +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = 'By-item hierarchy coefficients') +
  xlab('')
```

Plot them. Next the hierarchy * gender interaction coefficients:

```{r, fig.width = 12, fig.height = 6}
interact_coefs %>% ggplot(aes(x = reorder(Pair, InteractionCoef),
                            y = InteractionCoef)) +
  geom_col(fill = 'peachpuff3') + theme_bham_stats +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = 'By-item hierarchy * gender coefficients') +
  xlab('')
```

## Merge this with corpus data

Load the corpus data:

```{r, message = FALSE}
items <- read_csv('../data/item_corpus.csv')
```

Create the log-transformed ratio measure:

```{r}
items <- mutate(items,
                LogMF_Power = log10((PowerfulM + 1) / (PowerfulF + 1)),
                LogMF_Powerless = log10((PowerlessM + 1) / (PowerlessF + 1)),
                LogRatio = LogMF_Power / LogMF_Powerless)
```


Rename the pair clumn so that it matches the coefficient data frames:

```{r}
items <- rename(items,
                Pair = PairName)
```

Merge:

```{r}
gender_coefs <- left_join(gender_coefs, items)
hierarchy_coefs <- left_join(hierarchy_coefs, items)
interact_coefs <- left_join(interact_coefs, items)
```

## Correlations, ratio difference (M/F ratio of powerful - M/F ratio of powerless):

Look at correlations, with by-participant hierarchy slopes with the ratio difference score:

```{r, fig.width = 8, fig.height = 6}
hierarchy_coefs %>% 
  ggplot(aes(x = RatioDiff,
                            y = HierarchyCoef)) +
  geom_point(size = 2) +
  geom_smooth(method = 'lm') +
  theme_bham_stats
```

Next, correlations with by-participant gender slopes:

```{r, fig.width = 8, fig.height = 6}
gender_coefs %>% 
  ggplot(aes(x = RatioDiff,
                            y = GenderCoef)) +
  geom_point(size = 2) +
  geom_smooth(method = 'lm') +
  theme_bham_stats
```

Finally, interaction:

```{r, fig.width = 8, fig.height = 6}
interact_coefs %>% 
  ggplot(aes(x = RatioDiff,
                            y = InteractionCoef)) +
  geom_point(size = 2) +
  geom_smooth(method = 'lm') +
  theme_bham_stats
```

Get the correlation coefficients for the ratio difference score:

```{r}
with(gender_coefs, cor(RatioDiff, GenderCoef))
with(hierarchy_coefs, cor(RatioDiff, HierarchyCoef))
with(interact_coefs, cor(RatioDiff, InteractionCoef))
```

## Correlations, ratio of ratis (M/F ratio of powerful / M/F ratio of powerless):

Look at correlations, with by-participant hierarchy slopes with the ratio of ratio score. Since there's a really extreme ratio, for plotting purposes this is excluded:

```{r, fig.width = 8, fig.height = 6}
hierarchy_coefs %>% 
  filter(DoubleRatio < 200) %>% 
  ggplot(aes(x = DoubleRatio,
                            y = HierarchyCoef)) +
  geom_point(size = 2) +
  geom_smooth(method = 'lm') +
  theme_bham_stats
```

Next, correlations with by-participant gender slopes:

```{r, fig.width = 8, fig.height = 6}
gender_coefs %>% 
  filter(DoubleRatio < 200) %>% 
  ggplot(aes(x = DoubleRatio,
                            y = GenderCoef)) +
  geom_point(size = 2) +
  geom_smooth(method = 'lm') +
  theme_bham_stats
```

Finally, interaction, and make this a publication-ready plot:

```{r, fig.width = 8, fig.height = 6}
interact_coefs %>% 
  mutate(Pair = str_replace_all(Pair, '_', ' '),
         Pair = str_replace_all(Pair, ':', '/')) %>% 
  filter(DoubleRatio < 200) %>% 
  ggplot(aes(x = DoubleRatio,
                            y = InteractionCoef,
             label = Pair)) +
  coord_cartesian(xlim = c(-4, 22)) +
  geom_smooth(method = 'lm', col = 'black', fill = 'peachpuff3') +
  geom_point(col = 'red', size = 2) +
  geom_text_repel(size = 4, force = 11) +
  geom_vline(xintercept = 1, linetype = 2, size = 2) +
  theme_bham_stats +
  labs(title = 'By-item Hierarchy * Gender interaction coefficients\nas a function of corpus-based language bias\n') +
  ylab('Hierarchy * Gender coefficient\n') +
  xlab('\nGender asymmetry: Male/female ratio for powerful divided by\nmale/female ratio for powerless') +
  theme(plot.title = element_text(face = 'bold', size = 20, hjust = 0.5),
        axis.title = element_text(face = 'bold', size = 20),
        axis.text.x = element_text(face = 'bold', size = 16),
        axis.text.y = element_text(face = 'bold', size = 12),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank())
ggsave('../figures/figure4.png', width = 14, height = 8)
```

## Perform correlations with random slopes

Get the correlation coefficients for the ratio difference score. With the high-leverage point:

```{r}
with(gender_coefs, cor(DoubleRatio, GenderCoef))
with(hierarchy_coefs, cor(DoubleRatio, HierarchyCoef))
with(interact_coefs, cor(DoubleRatio, InteractionCoef))
```

... without the high-leverage point:

```{r}
with(filter(gender_coefs, DoubleRatio < 200), cor(DoubleRatio, GenderCoef))
with(filter(hierarchy_coefs, DoubleRatio < 200), cor(DoubleRatio, HierarchyCoef))
with(filter(interact_coefs, DoubleRatio < 200), cor(DoubleRatio, InteractionCoef))
```

## Estimate uncertainty in these correlations with Bayesian regression models:

Z-transform both predictor and response to make the regression model equal to a correlation. Also run this without the influential point (similar results are obtained if it's included, as the correlations above show).

```{r}
gender_z <- gender_coefs %>% 
  # filter(DoubleRatio < 200) %>% 
  mutate(Ratio_z = DoubleRatio - mean(DoubleRatio),
         Ratio_z = Ratio_z / sd(Ratio_z),
         Gender_z = GenderCoef - mean(GenderCoef),
         Gender_z = Gender_z / sd(Gender_z))

hierarchy_z <- hierarchy_coefs %>% 
  # filter(DoubleRatio < 200) %>% 
  mutate(Ratio_z = DoubleRatio - mean(DoubleRatio),
         Ratio_z = Ratio_z / sd(Ratio_z),
         Hierarchy_z = HierarchyCoef - mean(HierarchyCoef),
         Hierarchy_z = Hierarchy_z / sd(Hierarchy_z))

interact_z <- interact_coefs %>% 
  # filter(DoubleRatio < 200) %>% 
  mutate(Ratio_z = DoubleRatio - mean(DoubleRatio),
         Ratio_z = Ratio_z / sd(Ratio_z),
         Inter_z = InteractionCoef - mean(InteractionCoef),
         Inter_z = Inter_z / sd(Inter_z))
```

Set the prior:

```{r}
my_priors <- c(prior(normal(0, 0.5), class = b))
```

Run the Bayesian regressions for correlations (without intercept since everything's centered).

For this analysis, the extremely high double ratio coefficient is dropped. The main result (hierarchy * condition coefficient ~ asymmetry correlation) is supported regardless of whether this value is included or not.

```{r, message = FALSE}
gender_brm <- brm(Gender_z ~ -1 + Ratio_z,
                  data = filter(gender_z, DoubleRatio < 200), prior = my_priors,
                  init = 0, seed = 42)
```

Hierarchy correlation:

```{r, message = FALSE}
hierarchy_brm <- brm(Hierarchy_z ~ -1 + Ratio_z,
                  data = filter(hierarchy_z, DoubleRatio < 200), prior = my_priors,
                  init = 0, seed = 42)
```

Interaction correlation:

```{r, message = FALSE}
interact_brm <- brm(Inter_z ~ -1 + Ratio_z,
                  data = interact_z, prior = my_priors,
                  init = 0, seed = 42)
```

Look at the output:

```{r}
summary(gender_brm)
summary(hierarchy_brm)
summary(interact_brm)
```

Marginal effects:

```{r}
marginal_effects(interact_brm, 'Ratio_z')
```

Posterior probability:

```{r}
sum(posterior_samples(interact_brm)$b_Ratio_z < 0) / nrow(posterior_samples(interact_brm))
```


This completes this analysis.


